### 2015-11-11 paper/paper.tex
puliendo

\documentclass[a4,useAMS,usenatbib,usegraphicx]{mn2e} 
%\documentclass{latex/emulateapj} 
%External Packages and personalized macros
%=========================================================================
%		EXTERNAL PACKAGES
%=========================================================================
\usepackage{amsmath} 
\usepackage{amssymb} 
%\usepackage[section]{placeins}
\usepackage {graphicx}
%\usepackage{graphics}
\usepackage[dvips]{epsfig}
\usepackage{epsfig}  
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{caption}
%Non reposionated tables
\usepackage{float}
\restylefloat{table}
%Multiple columns support for tables
\usepackage{array}
\usepackage{booktabs}
\setlength{\heavyrulewidth}{1.5pt}
\setlength{\abovetopsep}{4pt}
\pdfminorversion=5

%=========================================================================
%		INTERNAL MACROS
%=========================================================================
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}

% To highlight comments 
\definecolor{red}{rgb}{1,0.0,0.0}
\newcommand{\red}{\color{red}}
\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
\newcommand{\SRK}[1]{\textcolor{darkgreen}{\bf SRK: \textit{#1}}}
\newcommand{\SRKED}[1]{\textcolor{darkgreen}{\bf #1}}
\newcommand{\before}[1]{\textcolor{red}{ #1}}
\newcommand{\after}[1]{\textcolor{darkgreen}{ #1}}

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\jcap}{JCAP}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc}\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc}\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
%SAMPLES
\newcommand{\GHBDM}{\texttt{GH}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\GHFOF}{\texttt{GH}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\IHBDM}{\texttt{IH}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\IHFOF}{\texttt{IH}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\PBDM}{\texttt{P}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\PFOF}{\texttt{P}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\IPBDM}{\texttt{IP}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\IPFOF}{\texttt{IP}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\RIPBDM}{\texttt{RIP}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\RIPFOF}{\texttt{RIP}$_{\mbox{\tiny{FOF}}}$ }

%MY COMMANDS #############################################################
\newcommand{\sub}[1]{\mbox{\scriptsize{#1}}}
\newcommand{\dtot}[2]{ \frac{ d #1 }{d #2} }
\newcommand{\dpar}[2]{ \frac{ \partial #1 }{\partial #2} }
\newcommand{\pr}[1]{ \left( #1 \right) }
\newcommand{\corc}[1]{ \left[ #1 \right] }
\newcommand{\lla}[1]{ \left\{ #1 \right\} }
\newcommand{\bds}[1]{\boldsymbol{ #1 }}
\newcommand{\oiint}{\displaystyle\bigcirc\!\!\!\!\!\!\!\!\int\!\!\!\!\!\int}
\newcommand{\mathsize}[2]{\mbox{\fontsize{#1}{#1}\selectfont $#2$}}
\newcommand{\eq}[2]{\begin{equation} \label{eq:#1} #2 \end{equation}}
\newcommand{\lth}{$\lambda_{th}$ }
\newcommand{\reff}{{\ifmmode{r_{\mbox{\tiny eff}}}\else{$r_{\mbox{\tiny eff}}$}\fi}}
%#########################################################################

\begin{document}

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[Halo concentration from the mass profile]{Estimating dark matter
  halo concentration using the integrated mass profile} 
\author[C. Poveda, J.E. Forero-Romero, J.C. Mu\~noz-Cuartas]{
\parbox[t]{\textwidth}{\raggedright 
  C. Poveda \thanks{cn.poveda542@uniandes.edu.co}$^{1}$
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$
  J. C. Mu\~noz-Cuartas \thanks{xxxx@udea.edu.co}$^{2}$
}
\vspace*{6pt}\\
$^1$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^2$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia
}

\maketitle

\begin{abstract}
We present a new algorithm to estimate the concentration of N-body
dark matter halos.
The method uses the integrated mass profile to find the optimal
concentration value. 
Its main advantage is that it uses the full particle information
without any binning, making it realiable in  cases when low numerical
resolution becomes a limitation for other methods.  
Additionally, the algorithm is straightforward to implement.  
We test the performance of this method by estimating halo
concentration both on mock and N-body halos. 
We compare these results against two other methods:
maximum radial velocity measurements and radial particle binning to
fit the density profile. 
Tests on  the mock halos show that the accuracy of our method to recover input
concentrations varies with halo resolution, outperforming the other
two methods. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy, while for the maximum radial
velocity method it is on the order of $20\%$ 
and density profile methods fail at the $100\%$ level.
For halos with $2\times 10^4$ particles our method achieves a recovery
accuracy of $0.01\%$ while the velocity and density profile methods
achieve $0.1\%$ and $1\%$, respectively. 
We also measure the mass-concentration relationship on N-body
data. We find that within the probed mass range ($10^{12}\hMsun <
10^{14}\hMsun$) the three methods give consistent results within the
statistical uncertainties.   
We only find a small discrepancy at low masses, $M<10^{13}\hMsun$, where
our method yields lower median concentration values by $15\%-20\%$
compared to the velocity and density methods. 
From these results we believe that the new method deserves to be
included in a simulator's toolkit to probe the internal structure of
dark matter halos.
\end{abstract}

\begin{keywords}
Cosmology: theory - large-scale structure of Universe -
Methods: data analysis - numerical - N-body simulations
\end{keywords}

%*************************************************************************
\section{Introduction}
\label{sec:introduction}
%*************************************************************************
In the current structure formation paradigm the properties of galaxies
are coupled to the evolution of their dark matter (DM) hosting halo.
This dependence has motivated the detailed study of dark matter halo
properties such as its formation history and internal structure. 
In particular, the sizes and dynamics of galaxies are determined by
the DM distribution inside a halo.

The internal DM distribution is parameterized through the density
profile. 
In a first approximation this profile is spherically symmetric and
its density only dependens on the radial coordinate. 
One the most popular parameterizations for DM halo radial
density distribution is the Navarro-Frenk-White (NFW) profile
\citep{NFW}.   
If one is not interested in the very central region of the
dark matter halo (where galaxy formation takes place, and where the
effects of baryon physics on dark matter distribution is still an
unknown) one can also assume that this profile is universal.
\citep{Navarro2010}.  
 This profile is a double power law in radius, where the transition
break happens at the so-called scale radius, $r_s$.  
The ratio between the scale radius and the virial radius $R_v$ is
known as the concentration $c=R_v/r_s$.

The concentration of the NFW profile provides the conceptual
framework to study simulated dark matter halos as a function of
redshift and cosmological parameters.
In these studies 
\citep{Neto2007,Maccio2008,Duffy2008,Munoz2011,Prada2012,Ludlow2014} 
the results are summarized through the
mass-concentration relationship; that is the distribution of
concentration values at a fixed halo mass and redshift.
A big challenge in such numerical experiments is having a
reliable algorithm to estimate the concentration. 
Such an algorithm should provide unbiased results and must be robust
when applied to simulations of different numerical resolution.

There are two main algorithms to estimate the concentration parameter
of a N-body dark matter halo. 
The first method takes the halo particles and bins them into
logarithmic radii to estimate the density in each bin, then it 
proceeds to fit the density as a function of the radius to the NFW
profile.  
A second method uses an analytic property of the NFW profile
that relates the maximum of the ratio of the circular velocity to the
virial velocity.  The concentration can be then found as the root of a
algebraic equation dependent on this maximum value.

The first method is straightforward to apply but presents two
disadvantages.  
First, it requires a large number of particles in
order to have a proper density estimate in each bin.  
This makes the method robust only for halos with at least $10^3$ particles.  
The second problem is that there is not a way to estimate the optimal
radial bin size, different choices produce different results for the
concentration.

The second method solves the two problems mentioned above.  
It works with low particle numbers and does not involve data binning.  
However, it effectively takes into account only a single data point and
discards the behaviour of the ratio $V_{\rm circ}/V_{\rm vir}$ below
and above its maxima.  
Small fluctuations on the value of this maximum can yield large
perturbations on the estimated concentration parameter.

In this letter we present a third alternative based on fitting the
integrated mass profile.
This approach has two advantages with respect to the above mentioned
methods.  
It does not involve any data binning and does not throw away data
points. 
This results in a robust estimate even at low resolution/particle
numbers.   
Furthermore, since the method does not require any binning, there is
no need to tune numerical parameters, providing a new independent
method to estimate the concentration parameter. 
Here we provide a detailed description of this method and a
comparison against the two traditional methods by measuring its
performance both on mock and N-body halos.

\section{Basic properties of the NFW density profile}
\label{sec:basics}

Let us review first the basic properties of the NFW density profile.
This shall help us to define our notation.

\subsection{Density profile}

The NFW density profile can be written as

\begin{equation}
\rho(r) = \frac{\rho_c\delta_c}{r/r_s(1+r/r_s)^2},
\label{eq:definition}
\end{equation}
%
where $\rho_c\equiv 3H^2/8\pi G$ is the Universe critical density, $H$
is the Hubble constant, $G$ and the universal gravitational constant,
$\delta_c$ is the halo dimensionless characteristic density and $r_s$
is the scale radius. 
This radius marks the point where the logarithmic slope of the density
profile is equal to -2, the transition between the power law
scaling $\rho\propto r^{-1}$ for $r<r_s$ and $\rho\propto r^{-3}$ for
$r>r_s$.

We define the virial radius of a halo, $r_v$, as the boundary of the
spherical volume that encloses a density of $\Delta_h$ times the mean
density of the Universe.  
The corresponding mass $M_{v}$, the virial mass, can be written as
$M_{v} = \frac{4\pi}{3}\bar{\rho}\Delta_h r_v^3$.  
From these virial quantities we define new dimensionless variables for
the radius and mass $x\equiv r/r_v$ and $m\equiv M(<r)/M_v$.

In this letter we use $\Delta_h=740$, a number roughly corresponding to
200 times the critical density at redshift z=0.

\subsection{Integrated mass profile}

From these definitions we can compute the total mass enclosed inside a
radius $r$:
\begin{equation}
M(<r) = 4\pi\rho_c\delta_c  r_s^3\left[\ln \left
  (\frac{r_s+r}{r_s}\right) - \frac{r}{r_s+r}\right],
\end{equation}
%
or in terms of the dimensionless mass and radius variables

\begin{equation}
m(<x) =
\frac{1}{A}\left[\ln\left(1+xc\right)-\left(\frac{xc}{xc+1}\right)\right],
\label{eq:profile}
\end{equation}
%
where
%
\begin{equation}
A=\ln\left(1+c\right)-\left(\frac{c}{c+1}\right),
\end{equation}
%
and the parameter $c$ corresponds to the concentration $c\equiv r_v/r_s$.

From this normalization and for later convenience we define the
following function
%
\begin{equation}
f(x) = \ln\left(1+x\right)-\left(\frac{x}{x+1}\right).
\label{eq:f_NFW}
\end{equation}
%
The most interesting feature of Eq. (\ref{eq:profile}) is that the
concentration is the only free parameter to describe the integrated
mass profile.

\subsection{Circular velocity profile}

It is also customary to express the mass of the halo in terms of the
circular velocity $V_{c}=\sqrt{GM(<r)/r}$.
From this we can define a new dimensionless circular velocity $v(<x)\equiv
V_{c}(<r)/V_{c}(<r_v)$, using the result in Eq. \ref{eq:profile}
to have:

\begin{equation}
v(<x)=\sqrt{\frac{1}{A}\left[\frac{\ln\left(1+xc\right)}{x}-\frac{c}{xc+1}\right]},
\end{equation}
%
This normalized profile always shows a maximum provided that the
concentration is larger than $c>2$.
It is possible to show that for the NFW profile the maximum is
provided by

\begin{equation}
\mathrm{max}(v(<x)) = \sqrt{\frac{c}{x_{\rm max}}\frac{f(x_{\rm
      max})}{f(c)}},
\label{eq:max_v}
\end{equation}
where $x_{\rm max}=2.163$ \citep{Klypin2014} and the function $f(x)$
correspondes to the definition in Eq. (\ref{eq:f_NFW}).

\section{Methods to estimate the concentration from N-body simulations}
\label{sec:method}

\subsection{Estimates from the density and velocity profiles}

To date, there are two standard methods to estimate concentrations in
dark matter halos extracted from N-body simulations.

The first method tries to directly estimate the density profile.  
It takes all the particles in the halo and bins them in the logarithm of
the radial coordinate from the halo center.  
Then, it estimates the density in each logarithmic bin counting the
particles and dividing by the corresponding shell volume.  
At this point is possible to make a direct fit to the density as a
function of the radial coordinate. 
This method has been broadly used for more than two decades to study
the mass-concentration-redshift relation of dark matter halos.

A second method uses the circular velocity profile.  
It finds the value of $x$ for which the normalized circular velocity
$v(<x)$ shows a maximum.  
Using this value it solves numerically for the corresponding value of
the concentration using Eq. (\ref{eq:max_v}). 
This method has been most recently used by \cite{Klypin2014} to study
the mass-concentration-redshift relation using the Multidark
Simulation Suite.

\subsection{Our proposal: estimate from the integrated mass profile}

Our method uses the integrated mass profile defined in
Eq. (\ref{eq:profile}).
We build it from N-body data following the next steps.
First, we define the center of the halo to be at the position of the
particle with the lowest gravitational potential.  
Then we rank the particles by their increasing radial distance from
the center. 
From this ranked list of $i=1,N$ particles, the total mass at a radius
$r_i$ is $M_i=i\times m_p$, where $r_i$ is the position of the $i$-th
particle and $m_p$ is the mass of a single computational particle. 
In this process we discard the particle at the center.

We divide the enclosed mass mass $M_i$ and the radii $r_i$ by their
virial values to obtain the dimensionless variables $m_i$ and $x_i$.
Once the mass profile is expressed in dimensionless variables the
concentration is the only free parameter. 
We then use an Affine Invariant Markov Chain Monte Carlo implemented in the
python module {\em emcee} \citep{emcee} to sample the likelihood
function distribution defined by ${\cal L}(c)\propto
\exp(-\chi^2(c)/2)$ where the $\chi^2(c)$ is 
written as

\begin{equation}
\chi^2(c)= \sum_{i=1}^{N}[\log m_i - \log m(< x_i;c)]^2,
\end{equation}
%
where $m(<x_i;c)$ corresponds to the values in Eq.(\ref{eq:profile})
at $x=x_i$ for a given value of the concentration parameter $c$ and
the $i$ index sums over all the particles in the numerical profile.
From the $\chi^2$ distribution we find the optimal value of the
concentration and its associated uncertainty.

\section{Results}
\label{sec:results}

We apply the three methods mentioned above on two different halo
samples.
The first sample is composed by mock halos generated to have known
concentration values in perfect spherical symmetry following an NFW
profile.    
Using this sample we want to quantify the ability to recover the
expected concentration as a function of particle numbe.
The second sample comes from a publicly available N-body cosmological
simulation.  
In this way we test the new method on more realistic and complex
data. 
We summarize the results on this dataset by showing the impact on the
mass-concentration relationship.

\subsection{Mock halo sample}

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{error.pdf}
\end{center}
\caption{Average value of the relative error in the concentration
  estimate, $\avg{|D|}$, as a function of the particle number $N$ in
  the set of mock halos. Different symbols represent different
  methods. Our method provide the most accurate estimate at fixed
  particle number $N$.  {\it Con respecto a esta figura tengo una
    pregunta. Yo ajusté los perfiles con menos de 200 partulas? no
    recuerdo!}
    \label{fig:error}}
\end{figure}

\label{sec:data}
\begin{figure*}
  \begin{center}
    \includegraphics[width=0.32\textwidth]{conc_mass_vel.pdf}
    \includegraphics[width=0.32\textwidth]{conc_mass_dens.pdf}
    \includegraphics[width=0.32\textwidth]{conc_dens_vel.pdf}
  \end{center}
  \caption{Comparison between the concentrations measured by our
    method and the maximum velocity (left) and density (middle)
    methods. The line indicates the equal value between the two
    techniques. The right panel compares the results of the maximum
    velocity and density methods.
  \label{fig:mdv}}
\end{figure*}

The method we use to generate the mock halos is based on the
integrated mass profile.  
We start by fixing the desired concentration $c$ and total number of
particles $N$ in the mock halo.  
With these values we define the mass element as $\delta m = 1/M$, corresponding
to the mass of each particle such that the total mass is one.   
Then for each particle, $i=1,\ldots,N$, we find the value of $r_i$
such that the difference 
% 
\begin{equation}
m(<r_i;c) - i \cdot \delta m
\end{equation}
%
is zero using Ridders' method.

The value of $r_i$ is the radius of the $i$-th particle of the mock
halo.  
Then we generate random polar and azimuthal angles $\theta$ and
$\phi$ for each particle to ensure spherical symmetry.  
Finally these three spherical coordinates are transformed into
Cartesian coordinates $(r,\theta,\phi) \rightarrow (x,y,z)$.

We generate in total $400$ mock halos split into four different groups
of $100$ halos each.  The four groups differ in the total number of
particles for their halos: $20$, $200$, $2000$ and $20000$.  
Inside each group the halos have random concentration values in the range
$1<c<20$ with a uniform distribution.  
For all these halos we find  the concentration values using the
density, velocity and integrated mass methods described in the
previous section.

We quantify the difference between the expected $c_{in}$ and measured
$c_{out}$ values by 
%
\begin{equation}
D=(c_{in}-c_{out})/c_{in},
\label{eq:D}
\end{equation}
%
and
%
\begin{equation}
\avg{|D|}=\frac{1}{\left|{\cal{H}}\right|}\sum_{{\cal{H}}} |D|,
\end{equation}
%
where ${\cal{H}}$ corresponds to a set of mock halos, and
$\left|{\cal{H}}\right|$ is the number of haloes in ${\cal{H}}$. 
  The larger the value of $\avg{|D|}$, the larger the discrepace of
  the concentrations obtained for that method.

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{concentration.pdf}
\end{center}
\caption{Mass-concentration relationship for the three different
  methods used on the same cosmological N-body data. 
  The lines correspond to the median concentration values in each bin.
  \label{fig:concentration}}
\end{figure}

Figure \ref{fig:error} shows the behaviour of $\avg{|D|}$ as a function of
halo particle number for the three different methods to estimate the
concentration.

At fixed particle numbers our method always shows the lowest
$\avg{|D|}$ values compared to the other two methods.  Its accuracy is
on the order of $10\%$ for $20$ particles in the halo, going down to
$0.01\%$ for halos with $20000$ particles.  The dependence of
$\avg{|D|}$ with the particle number $N$ goes approximately as
$\avg{|D|}\propto N^{-1/2}$, hinting that the accuracy of the method
is related to a decrease of Poisson noise.

The method based on the maximum of the circular velocity shows a
similar behaviour $\avg{|D|}\propto N^{-1/2}$.  Its accuracy is $2-5$
times less than in our method, on the order of $20\%$ for $20$
particle halos and $0.5\%$ for $20000$ particle halos. {\bf The method
  based on the direct density fit shows the lowest accuracy for a low
  particle number. As it is discussed in \citep{Munoz2011}, density
  binning for halos with particle numbers below $\sim200$ leads to a
  biased estimation of the mass density profile, and therefore to a
  biased estimation of the concentration parameter. This behaviour is
  evident in the large values of $\avg{|D|}$ for halos with number of
  particles below ~200 and an intermediate accuracy between the other
  two methods for a high particle number.}

\subsection{Halos from a cosmological simulation}

{\bf The results presented in the previous section show that in an
  idealized setup (perfect spherical halos, pure NFW desity profiles,
  total isolation) our method of the mass profile performs better than
  the other two methods commonly used to measure concentrations in
  simulated dark matter haos. However, actual simulated dark matter
  halos are much more complex. They are not isolated, many of them
  experience mergers that take them out of the equilibrium. Halos may
  have also plenty of substructure, etc. All these effects disturb the
  mass density profile, taking it away from the ideal NFW profile we
  assume in this wok. In order to see the real performance of our new
  method use it to estimate the concentration parameter in dark matter
  halos obtained from a cosmological Nbody simulation and compare with
  the other two calsical methods.}

We use data from the MultiDark cosmological simulation that follows
the non-linear evolution of a dark matter density field sampled with
$2048^3$ particles over a cubic box of $1000\ \hMpc$ on a side.  The
data is publicly available through \url{http://www.cosmosim.org/}.
More details about the structure of the database and the simulation
can be found in \citep{2013AN....334..691R}.

We build a sample of all halos located in a cubic sub-volume of $100$
\hMpc on a side centered on the most massive halo in the simulation at
$z=0$ which corresponds to the \texttt{miniMDR1} tables in the
database.  From this sample we select all the halos at $z=0$ detected
with a Friends-of-Friends (FoF) algorithm with masses in the interval
$10^{12}\leq M_{\rm FoF}/\hMsun \leq 10^{15}$.  The FoF algorithm used
with a linking length of $0.17$ times the {\bf mean} inter-particle
distance. This choice translates into an overdensity $\Delta_h\sim
400-700$ dependent on the halo concentration \citep{More2011}.
Finally, we select from the database all the particles that belong to
each FoF group.

From this set of particles we follow the procedure spelled out in
Section \ref{sec:method} with $\Delta_h=740$ (corresponding to $200$
times the critical density) to select an spherical region that we
redefine to be our halo.  This choice makes that our overdensities are
fully included inside the original FoF particle group.  On the
interest of providing a fair comparison against the density method we
only report results from overdensities with at least $200$ particles.

Figure \ref{fig:concentration} shows the results of the
mass-concentration relationship.

%{\em Aqui hay algo importante que hay que entender antes de discutir
%  esta gráfica , y es la razón por la cual el metodo del perfil de
%  masa da valores de concentración tan bajos a baja masa y comparables
%  a los de los otros dos métodos a alta masa. Es importante entender
%  la razón de esto!.  Se me ocurren unas preguntas para tratar de
%  entender. \\ 1) La grafica de c-M \ref{fig:concentration} usa todos
%  los halos o solo los halos relajados? La inclusión de halos no
%  relajados introduce in bias en la mediana en Cvir, así que vale la
%  pena inclusive agregar una linea mas a la gráfica con los halos
%  relajados y sin ellos. Para los otros dos métodos ya se sabe que
%  hace la población de halos no relajados en la medida así que no creo
%  necesario agregarlas para esos métodos. \\ Hay que explicar un poco
%  mas la gráfica en el sentido de que se grafica y como se
%  calculó. mean concentration?, Median concentrations? o mean in
%  logscale?  Mas bins de cuanto? Mínimo cuantos halos por mass bin?,
%  Halos relajados en todos los casos, o indistintos?}

\section{Conclusions}
\label{sec:conclusions}

In this letter we presented a new method to estimate the concentration
of dark matter halos in N-body simulations.  
We tested our method on mock halo data to study the impact of total
number of particles and input concentration on the retrieved values.  
We compared these
results against two other methods commonly used in the literature to
estimate concentrations.  Finally, we applied our method to halos
extracted from a cosmological N-body simulation to estimate the impact
of our method on the mass concentration relationship.

The first benchmark was performed on mock halos generated with known
concentration values for different particle numbers.  For all methods,
the accuracy in retrieveing the input concentration increases with the
number of particles as summarized in Figure \ref{fig:error}.  Our
method systematically shows smaller errors in the retrieved values
compared to the other two methods, going from $10\%$ for halos with
$20$ particles down to $0.01\%$ for halos with $2\times 10^4$
particles.  In our method the average error on the retrieved values
decreases with the number of particles as $\langle|D|\rangle\propto
N^{-1}$.  The error in density method initially behaves as
$\langle|D|\rangle\propto N^{-3/2}$ but at $N=2\times 10^3$ it
saturates to $0.1\%$.  For the velocity method the error behaves
closely to $\langle|D|\rangle\propto N^{-1/2}$.  We also gauged the
effect of different concentration parameters at fixed particle
numbers.  We found larger offsets for larger input concentrations.
The offset goes in the sense that the measured concentration tends to
be underestimated.  Nevertheless, the dominant effect on the
measurement is the particle number.

We used a N-body cosmological simulation to test the impact of our
method on the mass-concentration relationship.  Although the three
methods are in general agreement there are noticeable differences in
the mass concentration relationship.

\bibliographystyle{mn2e}
\bibliography{references}

\end{document}